---
title: "Full interactomes and interspecies interactions"
author: "Vitalii Kleshchevnikov"
date: "17/10/2017"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install MItools

```{r}
if(!"MItools" %in% installed.packages()){
  devtools::install_github("vitkl/MItools")
}
suppressPackageStartupMessages({
  library(MItools)
})
```

## Load full human-human interactome and count interactions

By species search is taxonomy-hierarchy aware.

```{r}
human = fullInteractome(taxid = 9606, database = "IntActFTP",
                        format = "tab25", clean = TRUE,
                        protein_only = TRUE,
                        directory = NULL) # keep data files inside R library - default
```

## Load all human-viral interactions

10239 is the top viral clade identifier. 

```{r}
human_viral = interSpeciesInteractome(taxid1 = 9606,  taxid2 = 10239,
                                      database = "IntActFTP", format = "tab25",
                                      clean = TRUE, protein_only = TRUE)
```

## Load all human-mouse interactions

```{r}
human_mouse = interSpeciesInteractome(taxid1 = 9606,  taxid2 = 10090,
                                      database = "IntActFTP", format = "tab25",
                                      clean = TRUE, protein_only = TRUE)
```

# Count interactors and interactions

```{r}
NuniqueInteractions(human)
NuniqueInteractors(human)

NuniqueInteractions(human_viral)
NuniqueInteractors(human_viral)

NuniqueInteractions(human_mouse)
NuniqueInteractors(human_mouse)
```

# filter human-human data by detection method and recount interactors and interactions

```{r}
# subset two-hybrid interactions
human_two_hybrid = subsetMITABbyMethod(MITABdata = human,
               Interaction_detection_methods = "MI:0018")
NuniqueInteractions(human_two_hybrid)
NuniqueInteractors(human_two_hybrid)

# subset all interactions but two-hybrid
human_NOT_two_hybrid = subsetMITABbyMethod(MITABdata = human,
                   Interaction_detection_methods = "MI:0018", inverse_filter = T)
NuniqueInteractions(human_NOT_two_hybrid)
NuniqueInteractors(human_NOT_two_hybrid)

# subset affinity purification - mass spectrometry interactions
human_AP_MS = subsetMITABbyMethod(MITABdata = human,
               Interaction_detection_methods = "MI:0004",  Identification_method = "MI:0433")
NuniqueInteractions(human_AP_MS)
NuniqueInteractors(human_AP_MS)
```

# filter human-human data by PMID and recount interactors and interactions

```{r}
# subset both published and unpublished Vidal group data
Vidal_all = subsetMITABbyPMIDs(MITABdata = human,
               PMIDs = c("25416956", "unassigned1304"))
NuniqueInteractions(Vidal_all)
NuniqueInteractors(Vidal_all)

# subset Mattias Mann 2015 paper data
Mann = subsetMITABbyPMIDs(MITABdata = human,
               PMIDs = "26496610")
NuniqueInteractions(Mann)
NuniqueInteractors(Mann)
```

You can get help and more details on these functions (for example, how to find which molecular ontology terms correspond to which methods): ?subsetMITABbyMethod

# Find interactions between components of the mediator complex in the Vidal dataset

```{r}
mediator_complex_proteins = fread("http://www.uniprot.org/uniprot/?query=GO:0016592%20AND%20taxonomy:9606&format=tab&columns=id")
mediator_complex = subsetMITABbyID(Vidal_all,
                                   ID_seed = mediator_complex_proteins$Entry,
                                   within_seed = T, only_seed2nonseed = F)
NuniqueInteractions(mediator_complex)
NuniqueInteractors(mediator_complex)
```

# Find interactions of the components of the mediator complex with other proteins

```{r}
mediator_complex_interactions = subsetMITABbyID(Vidal_all,
                                   ID_seed = mediator_complex_proteins$Entry,
                                   within_seed = F, only_seed2nonseed = T)
NuniqueInteractions(mediator_complex_interactions)
NuniqueInteractors(mediator_complex_interactions)
```
