---
title: "Getting started: full interactomes and interspecies interactions"
author: "Vitalii Kleshchevnikov"
date: "17/10/2017"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This article shows how to install MItools package and download protein interaction data for one species (for example, human), 2 species/taxomony groups (for example, human and viruses) and how to filter this data by interaction detection method, publication ID (PMID) or list of protein IDs (UniprotKB).  

## Install MItools

```{r}
if(!"MItools" %in% installed.packages()){
  devtools::install_github("vitkl/MItools", dependencies = TRUE)
}
suppressPackageStartupMessages({
  library(MItools)
})
```

## Load full human-human interactome

Interactome search by species is taxonomy-hierarchy aware.

```{r}
human = fullInteractome(taxid = 9606, database = "IntActFTP",
                        format = "tab25", clean = TRUE,
                        protein_only = TRUE,
                        directory = NULL) # keep data files inside R library - default
```

## Load all human-viral interactions

10239 is the top viral clade identifier. 

```{r}
human_viral = interSpeciesInteractome(taxid1 = 9606,  taxid2 = 10239,
                                      database = "IntActFTP", format = "tab25",
                                      clean = TRUE, protein_only = TRUE)
```

## Load all human-mouse interactions

```{r}
human_mouse = interSpeciesInteractome(taxid1 = 9606,  taxid2 = 10090,
                                      database = "IntActFTP", format = "tab25",
                                      clean = TRUE, protein_only = TRUE)
```

# Count interactors and interactions

```{r}
uniqueNinteractions(human)
uniqueNinteractors(human)

uniqueNinteractions(human_viral)
uniqueNinteractors(human_viral)

uniqueNinteractions(human_mouse)
uniqueNinteractors(human_mouse)
```

# filter human-human data by detection method and recount interactors and interactions

```{r}
# subset two-hybrid interactions
human_two_hybrid = subsetMITABbyMethod(MITABdata = human,
               Interaction_detection_methods = "MI:0018")
uniqueNinteractions(human_two_hybrid)
uniqueNinteractors(human_two_hybrid)

# subset all interactions but two-hybrid
human_NOT_two_hybrid = subsetMITABbyMethod(MITABdata = human,
                   Interaction_detection_methods = "MI:0018", inverse_filter = T)
uniqueNinteractions(human_NOT_two_hybrid)
uniqueNinteractors(human_NOT_two_hybrid)

# subset affinity purification - mass spectrometry interactions
human_AP_MS = subsetMITABbyMethod(MITABdata = human,
               Interaction_detection_methods = "MI:0004",  Identification_method = "MI:0433")
uniqueNinteractions(human_AP_MS)
uniqueNinteractors(human_AP_MS)
```

# filter human-human data by PMID and recount interactors and interactions

```{r}
# subset both published and unpublished Vidal group data
Vidal_all = subsetMITABbyPMIDs(MITABdata = human,
               PMIDs = c("25416956", "unassigned1304"))
uniqueNinteractions(Vidal_all)
uniqueNinteractors(Vidal_all)

# subset Mattias Mann 2015 paper data
Mann = subsetMITABbyPMIDs(MITABdata = human,
               PMIDs = "26496610")
uniqueNinteractions(Mann)
uniqueNinteractors(Mann)
```

You can get help and more details on these functions (for example, how to find which molecular ontology terms correspond to which methods): ?subsetMITABbyMethod

# Find interactions between components of the mediator complex in the Vidal dataset

```{r}
mediator_complex_proteins = fread("http://www.uniprot.org/uniprot/?query=GO:0016592%20AND%20taxonomy:9606&format=tab&columns=id")
mediator_complex = subsetMITABbyID(Vidal_all,
                                   ID_seed = mediator_complex_proteins$Entry,
                                   within_seed = T, only_seed2nonseed = F)
uniqueNinteractions(mediator_complex)
uniqueNinteractors(mediator_complex)
```

# Find interactions of the components of the mediator complex with other proteins

```{r}
mediator_complex_interactions = subsetMITABbyID(Vidal_all,
                                   ID_seed = mediator_complex_proteins$Entry,
                                   within_seed = F, only_seed2nonseed = T)
uniqueNinteractions(mediator_complex_interactions)
uniqueNinteractors(mediator_complex_interactions)
```

# Date and packages used

```{r}
Sys.Date. = Sys.Date()
Sys.Date.
session_info. = devtools::session_info()
session_info.
```
