---
title: 'MItools vs igraph: calculating degree of a protein'
author: "Vitalii Kleshchevnikov"
date: "14/07/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages({
  library(MItools)
  library(data.table)
  library(igraph)
  library(microbenchmark)
})
```

This article shows how to download protein interaction data using MItools, import it into igraph and calculate degree of each protein using MItools and igraph.  

## Download full human interactome as a data.table using MItools

```{r}
human_interactome = fullInteractome(taxid = "9606", database = "IntActFTP",
                                    format = "tab27", clean = T, protein_only = TRUE)
```

## Convert data table to igraph object

```{r}
human_interactome_igraph = graph_from_data_frame(d = human_interactome$data[,
                             .(IDs_interactor_A, IDs_interactor_B,
                             Publication_Identifiers, Confidence_values)],
                             directed = F)
```

## Calculate node degree distribution using MItools

microbenchmark function from microbenchmark R package is the most convenient and accurate way to measure how long code runs.
Let's use this to find which, MItools or igraph will calculate node degree faster.

```{r}
microbenchmark({edgelist2degree(human_interactome$data)})
```

## Calculate node degree distribution using igraph

```{r}
# igraph is much faster than MItools
microbenchmark({degree(human_interactome_igraph)})
# but only if you don't account for time it takes to create an igraph object
microbenchmark({
  human_interactome_igraph = graph_from_data_frame(d = human_interactome$data[,
                             .(IDs_interactor_A, IDs_interactor_B,
                             Publication_Identifiers, Confidence_values)],
                             directed = F)
  degree(human_interactome_igraph)
  })
```

# Date and packages used

```{r}
Sys.Date. = Sys.Date()
Sys.Date.
session_info. = devtools::session_info()
session_info.
```

